\chapter{Methodology}\label{ch:methodology}

% ============================================================
% Cau truc hoc tu paper Q1/Q2:
%  - Method viet theo khoi "Step 1 ... Step k": Lam et al. (2023) muc 2.2-2.3; Saputro (2023)
%  - AHP crisp PHAI co bang Random Index + CR check: theo Lam & Gaur
%  - So lieu Likert/AHP: 03_phan-tich/tieu-chi/ va 03_phan-tich/ahp/ (sinh boi script)
% ============================================================

\section{The Proposed Research Model}\label{sec:framework}

The study is organised into three sequential stages, each of which answers one of the
research questions stated in \Cref{sec:objectives}. The first stage develops the set of
evaluation criteria: candidate criteria are synthesised from the literature reviewed in
\Cref{ch:literature}, refined through expert interviews, and then screened through a
large-scale survey, so that only criteria whose importance is confirmed by actual office
workers enter the quantitative model. The second stage quantifies the relative importance
of the retained criteria with AHP, including an explicit consistency check before the
weights are accepted. The third stage applies the model to a concrete purchase scenario:
candidate laptops are identified from the recommendation lists of professional review
outlets, their attribute data are collected from the Vietnamese retailer CellphoneS and
operationalised into a decision matrix, and TOPSIS ranks them using the weights obtained
in the second stage. The stages are
strictly ordered: the pairwise comparison in Stage~2 cannot be designed until the criteria
set is fixed, and the ranking in Stage~3 is meaningful only after the weights have passed
the consistency check. \Cref{fig:framework} summarises this pipeline.

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[
      node distance=7mm and 9mm,
      stage/.style={draw, rounded corners=2pt, align=center, minimum width=34mm,
                    minimum height=11mm, font=\small},
      substep/.style={draw, align=center, minimum width=30mm, minimum height=8mm,
                      font=\footnotesize},
      lbl/.style={font=\footnotesize\itshape},
      arr/.style={-{Stealth[length=2.5mm]}, thick}
    ]
    % Stage 1 column
    \node[substep] (lit) {Literature synthesis};
    \node[substep, below=of lit] (int) {Expert interviews\\ (purposive, $n=5$)};
    \node[substep, below=of int] (sur) {Likert survey screening\\ (retain if mean $\geq 3.0$)};
    \node[stage, below=of sur] (s1out) {Retained criteria};
    % Stage 2
    \node[stage, right=14mm of s1out] (s2) {Stage 2: AHP\\ weighting + CR check};
    % Stage 3
    \node[stage, right=14mm of s2] (s3) {Stage 3: Scenario\\ application + TOPSIS};
    % labels
    \node[lbl, above=2mm of lit] {Stage 1: Criteria development};
    \node[lbl, below=1mm of s2] {answers RQ2};
    \node[lbl, below=1mm of s3] {answers RQ3};
    \node[lbl, below=1mm of s1out] {answers RQ1};
    % arrows
    \draw[arr] (lit) -- (int);
    \draw[arr] (int) -- (sur);
    \draw[arr] (sur) -- (s1out);
    \draw[arr] (s1out) -- node[above, font=\footnotesize] {weights} (s2);
    \draw[arr] (s2) -- node[above, font=\footnotesize] {ranking} (s3);
  \end{tikzpicture}
  \caption{The proposed research model.}
  \label{fig:framework}
  \floatnote{\emph{Source:} authors' own elaboration.}
\end{figure}

\section{Stage 1 --- Criteria Development}\label{sec:stage1}

\subsection{Literature Synthesis}\label{subsec:lit-synthesis}

The starting point of the criteria set is the body of laptop-selection and
consumer-behaviour studies reviewed in \Cref{sec:prior-studies}. From these sources,
every attribute that at least one prior study used to evaluate laptops or closely related
devices was extracted and consolidated: price appears in nearly all of them
\parencite{sonmez_cakir_determination_2020,weng_siew_lam_evaluation_2023,liao_determinants_2022,rau_optimal_2018},
core hardware attributes such as processor, memory, and storage recur across the MCDM
applications \parencite{gaur_application_2023,maghsoudi_hybrid_2026}, and service-related
attributes such as brand and after-sales support are emphasised in the review-mining and
consumer-behaviour literature \parencite{kang_study_2022,sostar_assessment_2023}.
Attributes that overlap in meaning were merged into a single criterion, and attributes tied
to specialised use cases outside the office-worker persona, such as colour-accurate
displays for design work, were excluded at this stage. The remaining attributes were then
organised into thematic groups reflecting different aspects of the purchase decision, such
as core hardware capability or trust and after-sales value, so that the pairwise
comparisons of \Cref{sec:stage2} could later be structured hierarchically rather than as a
single flat list. The resulting set of candidate criteria is reported in
\Cref{subsec:criteria-set}.

\subsection{Expert Interviews}\label{subsec:expert-interviews}

Because the candidate criteria from \Cref{subsec:lit-synthesis} were derived from studies
conducted outside Vietnam, they were validated against local practice before being
quantified. Five experts
were selected purposively so that each occupies a different position in the laptop
procurement process: a deputy head of an international relations office with seven years
of end-user experience (CG-01), an IT administrator who specifies and maintains office
fleets (CG-02), an import--export documentation officer representing the target persona
(CG-03), a head of product development with ten years of managerial purchasing experience
(CG-04), and a retail sales consultant who advises laptop buyers daily (CG-05).

The interview protocol itself, rather than being improvised, was constructed following the
Interview Protocol Refinement (IPR) framework, whose four phases require that every
question be traceable to a research objective, that the questions be sequenced as an
inquiry-based conversation rather than a checklist, that the draft be reviewed before
fieldwork, and that it be piloted \parencite{castillo-montoya_preparing_2016}. Applying
this framework produced a five-part funnel that moves from the concrete to the general: a
warm-up section grounding each respondent in a recent, specific laptop-selection episode; a
stimulus-based section in which a literature finding for each criterion group is read
aloud immediately before the matching question, so that agreement or disagreement is
anchored to a stated claim rather than solicited as an abstract opinion; an open section
probing for criteria specific to the Vietnamese context that the literature synthesis
could not have anticipated; a trade-off scenario that generates qualitative seed data for
the pairwise comparisons of \Cref{sec:stage2}; and a closing section built around
member-checking, in which the interviewer restates the respondent's positions for
confirmation. Most substantive questions in this structure carry a follow-up probe that
asks the respondent to recall the specific case behind an answer or to explain the
reasoning process that produced it, a technique adapted from cognitive interviewing that
keeps a response anchored in the respondent's actual experience rather than a general
impression formed on the spot \parencite{willis_cognitive_2005}. This funnel structure and
probing technique operate inside a semi-structured session that functions as the
qualitative first round of a Delphi-style refinement preceding AHP, an arrangement applied
in prior criteria-development work \parencite{al_hazza_integrated_2022,mcmillan_how_2016}.
Each session lasted twenty to thirty minutes; all five respondents gave informed consent,
no audio was recorded, and only written notes were retained.

Interview notes were reviewed for proposals to add, remove, or redefine a criterion. To
keep the revision process auditable rather than ad hoc, a proposal was adopted only when it
was corroborated independently by a majority of the five experts, and any resulting change
to the candidate set was cross-checked against the literature reviewed in
\Cref{sec:prior-studies} before being finalised. The outcome of this revision process is
reported together with the resulting criteria set in \Cref{subsec:criteria-set}.

\subsection{Survey-Based Screening}\label{subsec:screening}

Expert validation establishes that a criterion is defensible, but not that the target
population actually weighs it when buying. The candidate criteria from
\Cref{subsec:expert-interviews} were therefore submitted to a wider screening survey
before any pairwise comparison was designed. A questionnaire was distributed through
Google Forms in which respondents rated the importance of each criterion when choosing a
laptop for office work on a five-point Likert scale, from 1 (not important) to 5 (very
important). The instrument deliberately contained one rating item per candidate criterion
and no other items, keeping completion time short and avoiding demographic questions that
the screening decision does not use.

The retention rule was fixed before data collection: a criterion is retained if its sample
mean is at least 3.0, the midpoint of the scale, and removed otherwise. On a five-point
scale the midpoint separates responses that lean towards importance from those that lean
away from it, so a mean below 3.0 indicates that respondents on balance do not regard the
criterion as important \parencite{chyung_evidencebased_2017}. Committing to the threshold
in advance prevents the rule from being adjusted after the results are known. Responses
missing more than 20\% of the items were to be discarded, and straight-lined responses,
in which a respondent gives the same rating to every item, were to be flagged for a group
decision rather than silently removed. The screening outcome, which reduces the candidate
set to the criteria carried into Stage~2, is reported in \Cref{sec:screening}.

\section{Stage 2 --- AHP for Criteria Weighting}\label{sec:stage2}

The relative importance of the criteria retained after screening is derived through the
Analytic Hierarchy Process, whose theoretical basis is reviewed in \Cref{sec:ahp-theory}.
The procedure follows five steps.

\paragraph{Step 1 --- Hierarchy construction.}
The decision problem is structured into three levels: the goal of selecting the most
suitable laptop for a Vietnamese office worker at the top, the criteria retained from
Stage~1 at the intermediate level, and the candidate laptop models at the bottom. Only
the upper two levels are involved in this stage; the alternatives enter the model in
Stage~3, where they are evaluated on directly measurable data rather than on pairwise
judgements.

\paragraph{Step 2 --- Pairwise comparison.}
The pairwise judgements were elicited from one representative expert drawn from the
interview panel of \Cref{subsec:expert-interviews}, selected for direct, first-hand
experience of office laptop procurement. For $n$ retained criteria the expert completes
$n(n-1)/2$ comparisons, each expressing how much more important one criterion is than
another on the nine-point ratio scale of \Cref{tab:saaty}. Relying on a single judgement
set concentrates the weighting in one perspective, which is why the consistency check in
Step~5 is applied strictly: an inconsistent matrix is returned to the expert for
re-evaluation rather than accepted or averaged away.

\begin{table}[htbp]
  \centering
  \caption{The Saaty fundamental scale of pairwise comparison.}
  \label{tab:saaty}
  \begin{tabular}{@{}cl@{}}
    \toprule
    \textbf{Intensity} & \textbf{Definition} \\
    \midrule
    1 & Equal importance \\
    3 & Moderate importance \\
    5 & Strong importance \\
    7 & Very strong importance \\
    9 & Extreme importance \\
    2, 4, 6, 8 & Intermediate values \\
    \bottomrule
  \end{tabular}
  \floatnote{\emph{Source:} adapted from \textcite{saaty_analytic_1987}.}
\end{table}

\paragraph{Step 3 --- Comparison matrix.}
For $n$ criteria, the pairwise judgements form a reciprocal matrix
$A = [a_{ij}]$ with $a_{ij} = 1/a_{ji}$ and $a_{ii} = 1$:
\begin{equation}\label{eq:pcm}
  A = \begin{bmatrix}
    1 & a_{12} & \cdots & a_{1n} \\
    1/a_{12} & 1 & \cdots & a_{2n} \\
    \vdots & \vdots & \ddots & \vdots \\
    1/a_{1n} & 1/a_{2n} & \cdots & 1
  \end{bmatrix}.
\end{equation}
Only the $n(n-1)/2$ entries above the diagonal are elicited; the remaining entries follow
from reciprocity, so the expert is never asked the same comparison twice.

\paragraph{Step 4 --- Priority (weight) vector.}
The criteria weights are taken as the principal right eigenvector of $A$, that is, the
vector $\mathbf{w}$ satisfying
\begin{equation}\label{eq:weight}
  A\mathbf{w} = \lambda_{\max}\mathbf{w}, \qquad \sum_{i=1}^{n} w_i = 1,
\end{equation}
where $\lambda_{\max}$ is the largest eigenvalue of $A$. Several simpler approximations
circulate in applied work, including the normalised column average used by
\textcite{weng_siew_lam_evaluation_2023}; the eigenvector is preferred here because
\textcite{saaty_decisionmaking_2003} shows it to be the theoretically required solution
whenever the judgement matrix is a small perturbation of a fully consistent one, which the
consistency check in Step~5 verifies, and because the comparative simulations of
\textcite{ishizaka_how_2006} find that approximation methods diverge from the eigenvector
increasingly as matrices grow larger or less consistent. In computation the eigenvector is
obtained by power iteration, repeatedly multiplying a normalised weight vector by $A$
until convergence. As a cross-check, the weights are also computed by the geometric mean
of each row; agreement between the two vectors in the ordering of criteria is treated as
evidence that the result does not depend on the derivation method.

\paragraph{Step 5 --- Consistency check.}
Because the judgements are elicited from a human respondent, they may contradict one
another, and the eigenvector of a heavily contradictory matrix is not a meaningful weight
vector. The internal coherence of the judgements is therefore verified through the
consistency ratio $CR = CI/RI$, where the consistency index is
\begin{equation}\label{eq:cr}
  CI = \frac{\lambda_{\max} - n}{\,n - 1\,},
\end{equation}
$\lambda_{\max}$ is the principal eigenvalue from \Cref{eq:weight}, and $RI$ is the random
index, the mean consistency index of randomly generated reciprocal matrices of the same
order (\Cref{tab:ri}). A matrix is accepted when $CR \le 0.10$
\parencite{Saaty2008DECISIONMW}. If the threshold is exceeded, the weights are not used:
the three pairwise entries that deviate most from the values a consistent matrix would
imply are identified and returned to the expert for re-evaluation, and the matrix is
recomputed. The threshold is never relaxed to accommodate a marginally failing matrix.

\begin{table}[htbp]
  \centering
  \caption{Random Index (RI) values by matrix order $n$.}
  \label{tab:ri}
  \begin{tabular}{@{}lccccccccc@{}}
    \toprule
    $n$ & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
    \midrule
    RI & 0.58 & 0.90 & 1.12 & 1.24 & 1.32 & 1.41 & 1.45 & 1.49 \\
    \bottomrule
  \end{tabular}
  \floatnote{\emph{Source:} adapted from \textcite{saaty_analytic_1987}.}
\end{table}

\section{Stage 3 --- Scenario-Based Application with TOPSIS}\label{sec:stage3}

\subsection{Decision Scenario}\label{subsec:scenario-design}

The third stage applies the weighted criteria to a concrete decision situation. The
scenario mirrors the problem statement of \Cref{sec:problem}: an office worker in Vietnam
needs a laptop for work, has browsed review websites and the catalogues of established
retailers, and still cannot decide, because the information available compares products
one attribute at a time while the actual decision requires weighing all attributes at
once. Framing Stage~3 as a scenario serves two purposes. It keeps the demonstration
honest, since every modelling choice, from which laptops enter the comparison to how each
criterion is measured, must be one this buyer could reproduce from publicly available
information; and it makes the output directly interpretable, since the resulting ranking
answers the question the buyer actually asks. The stage proceeds in three steps that
parallel a standard data workflow: identifying the alternatives, collecting and
operationalising their attribute data, and analysing the resulting decision matrix with
TOPSIS.

\subsection{Alternative Identification}\label{subsec:alternatives}

The candidate laptops are not chosen by the researchers' own judgement, which would risk
pre-selecting the winner, but harvested from published recommendation lists of
professional review outlets and then filtered for local availability. An outlet is
included in the source pool if it satisfies three conditions: it maintains a documented,
standardised laboratory testing protocol rather than opinion-only reviews; it actively
covers laptops as a product category; and its editorial operation is stable enough that
its current recommendations reflect that protocol. Seven outlets meet these conditions,
summarised in \Cref{tab:review-sources}. One widely known outlet, Engadget, was
considered and excluded: it reduced its editorial staff by roughly one third in 2024,
reoriented its output towards search-driven affiliate content, and changed ownership in
early 2026, so the stability condition is not met. CNET is retained with a caveat: the
outlet published AI-generated articles without clear disclosure in 2023, but it operates
its own testing laboratory and tightened human editorial review after the incident, and
its current laptop recommendations are lab-based.

\begin{table}[htbp]
  \centering
  \caption{Review outlets used as sources for alternative identification.}
  \label{tab:review-sources}
  \begin{tabular}{@{}lp{3.6cm}p{5.8cm}@{}}
    \toprule
    \textbf{Outlet} & \textbf{Operator} & \textbf{Testing approach} \\
    \midrule
    PCMag        & Ziff Davis          & Standardised lab benchmarks since 1982; Editors' Choice programme \\
    Tom's Guide  & Future plc          & Comparative product testing; seasonal category rankings \\
    TechRadar    & Future plc          & Global review team; per-segment rankings (business, budget) \\
    Laptop Mag   & Future plc          & Laptop-only coverage; in-house battery and performance benchmarks \\
    CNET         & Ziff Davis          & In-house lab for performance, battery, and display testing \\
    Wirecutter   & The New York Times  & Hands-on testing; no sponsored review placements \\
    RTINGS       & Independent         & Instrumented measurement (display, thermals, input latency) \\
    \bottomrule
  \end{tabular}
  \floatnote{\emph{Source:} authors' own compilation from the editorial policies of the
  listed outlets.}
\end{table}

From each outlet in the pool, the current recommendation list for the best business or
premium laptops of the model year is collected, and the union of these lists forms the raw
candidate set. Because the same model frequently appears on several lists, the set is
deduplicated by manufacturer and model designation, treating configuration variants of the
same model as a single candidate at this step. Price and configuration data for the
deduplicated set are then collected from Amazon.com rather than from a Vietnamese retailer,
because the outlets' recommendations are dominated by newly launched, internationally
released configurations that Vietnamese retail catalogues do not yet carry at the time of
collection, so the buyer of the scenario in \Cref{subsec:scenario-design} would in practice
have to source most of these models through an international retailer regardless of the
Vietnamese market context. Amazon.com is the world's largest online retailer and the
leading e-commerce platform in the United States, and using its product listings as a
source of laptop specification and price data has direct precedent in the
consumer-electronics literature: \textcite{rau_optimal_2018} construct their notebook
price-history dataset directly from Amazon product listings. The date of collection is
recorded with the data, since both the recommendation lists and the retailer's catalogue
change over time.

\subsection{Data Collection and Operationalisation}\label{subsec:operationalisation}

Each remaining alternative is scored on the retained criteria, and the way each
criterion is measured is fixed before the data are collected. Five criteria are directly
quantitative and are taken as listed on Amazon.com: the price in US dollars, the memory
capacity in gigabytes, the storage capacity in gigabytes, the device weight in kilograms
(converted from the pounds figure shown in the listing), and the battery capacity in
watt-hours as specified by the manufacturer. Watt-hours are used in preference to a
manufacturer-claimed runtime in hours, because the latter is measured under differing,
vendor-specific workloads and is therefore not commensurable across brands, whereas
watt-hour capacity is a fixed physical specification of the battery itself. The processor
cannot be treated this way: chipset designations such as product names and generation
labels are neither numeric nor directly comparable across manufacturers. The criterion is
therefore operationalised through a standardised benchmark score, using the PassMark CPU
database, which converts heterogeneous processor names into a single performance scale.
This step matters for the validity of the analysis, because TOPSIS requires attribute
values that are numeric, monotonic in desirability, and commensurable
\parencite{behzadian_state--art_2012}; a benchmark score satisfies all three conditions
where a chipset name satisfies none. Brand, the one qualitative criterion, is likewise
converted into an observable quantity rather than scored by hand: it is operationalised as
each manufacturer's worldwide PC shipment market share for the first quarter of 2026, as
reported by the market-research firm Gartner \parencite{gartner_2026}. This proxy is
chosen in preference to a Vietnam-specific measure, such as a count of authorised service
centres, because no freely accessible source publishes brand-level market share for the
Vietnamese laptop market specifically; global shipment share is disclosed quarterly and is
independently reproducible, and it still functions as the concrete, verifiable expression
of manufacturer standing that the trust and after-sales group of \Cref{tab:criteria}
calls for, whereas a hand-assigned brand score would reintroduce exactly the subjectivity
that the screening and weighting stages were designed to control. Gartner is used in
preference to IDC, the other major PC-market tracker, because Gartner's quarterly release
reports a share figure for every one of the six brands present in this candidate set,
including Acer, whereas IDC's release names only its top five vendors by shipment volume
and groups every remaining brand, Acer included, into a single undifferentiated "other
vendors" residual; using a single source that individually reports all six brands avoids
mixing two trackers' figures within the same criterion column.

\subsection{The TOPSIS Procedure}\label{subsec:topsis-procedure}

Given the criteria weights from Stage~2 and the decision matrix assembled above, the
candidate laptops are ranked with TOPSIS over seven steps.

\paragraph{Step 1 --- Decision matrix.}
For $m$ alternatives and $n$ criteria, $X = [x_{ij}]$ collects the performance of
alternative $i$ on criterion $j$.

\paragraph{Step 2 --- Vector normalisation.}
\begin{equation}\label{eq:norm}
  r_{ij} = \frac{x_{ij}}{\sqrt{\sum_{i=1}^{m} x_{ij}^{2}}}.
\end{equation}

\paragraph{Step 3 --- Weighted normalised matrix.}
\begin{equation}\label{eq:weighted}
  v_{ij} = w_j \, r_{ij}.
\end{equation}

\paragraph{Step 4 --- Ideal solutions.}
The positive ideal solution $A^{+}$ and negative ideal solution $A^{-}$ are formed from
the best and worst weighted scores of each criterion, respecting its benefit or cost direction:
\begin{equation}\label{eq:pis-nis}
  A^{+} = \{v_1^{+},\dots,v_n^{+}\}, \qquad
  A^{-} = \{v_1^{-},\dots,v_n^{-}\}.
\end{equation}

\paragraph{Step 5 --- Separation measures.}
\begin{equation}\label{eq:sep}
  S_i^{+} = \sqrt{\sum_{j=1}^{n}\left(v_{ij}-v_j^{+}\right)^{2}}, \qquad
  S_i^{-} = \sqrt{\sum_{j=1}^{n}\left(v_{ij}-v_j^{-}\right)^{2}}.
\end{equation}

\paragraph{Step 6 --- Closeness coefficient.}
\begin{equation}\label{eq:closeness}
  C_i = \frac{S_i^{-}}{\,S_i^{+} + S_i^{-}\,}, \qquad 0 \le C_i \le 1.
\end{equation}

\paragraph{Step 7 --- Ranking.}
The alternatives are ranked in descending order of $C_i$; the laptop with the highest
$C_i$ is the recommended choice.
