\chapter{Results}\label{ch:results}

% ============================================================
% Nguyen tac: Results PHAN CHIEU 1-1 voi Methodology (mau Maghsoudi 2026, muc 4.x <-> 3.x).
% 4.1 (criteria dev outcomes) + 4.2 (screening) + 4.3 (AHP weights) DA VIET tu so lieu
% script chinh thuc:
%   03_phan-tich/tieu-chi/2026-07-11_likert-ket-qua_v1.0_final_user.txt
%   03_phan-tich/ahp/2026-07-11_ahp-ket-qua_v1.0_final_user.txt
% 4.4-4.6 cho Phase 5 (TOPSIS).
% 2026-07-12: tach §4.1 khoi Ch3 theo yeu cau user — Methodology (Ch3) chi mo ta
% PHUONG PHAP/quy trinh, khong nen ket qua ap dung phuong phap (IMRaD chuan).
% ============================================================

\section{Criteria Development Outcomes}\label{sec:criteria-outcome}

\subsection{The Candidate Criteria Set}\label{subsec:criteria-set}

Applying the literature-synthesis procedure of \Cref{subsec:lit-synthesis} to the studies
reviewed in \Cref{sec:prior-studies} produced fourteen candidate criteria. These were
organised into six groups reflecting the aspects of the purchase decision they describe:
economic cost, core hardware capability, design and interaction experience, portability
and durability, ecosystem utility, and trust and after-sales value. \Cref{tab:criteria}
lists the resulting set together with each criterion's group and its direction of
preference. Most criteria are benefit criteria, for which a larger value is preferred;
price and physical size or weight are cost criteria, for which a smaller value is
preferred. The distinction matters at two later points. It determines the direction of the
ideal solutions in TOPSIS (\Cref{sec:stage3}), and it guards against the error of treating
a heavier or more expensive machine as better simply because its raw attribute value is
larger.

\begin{table}[htbp]
  \centering
  \caption{The fourteen candidate evaluation criteria in six groups.}
  \label{tab:criteria}
  \begin{tabularx}{\textwidth}{@{}L{1.2cm}L{4.6cm}L{2.2cm}Y@{}}
    \toprule
    \textbf{ID} & \textbf{Criterion} & \textbf{Type} & \textbf{Group} \\
    \midrule
    C1  & Price                   & Cost    & Economic \\
    C2  & CPU                     & Benefit & Core hardware \\
    C3  & RAM                     & Benefit & Core hardware \\
    C4  & Storage (SSD)           & Benefit & Core hardware \\
    C5  & Display                 & Benefit & Design \& interaction \\
    C6  & Keyboard                & Benefit & Design \& interaction \\
    C7  & Design \& shell material & Benefit & Design \& interaction \\
    C8  & Size / weight           & Cost    & Portability \& durability \\
    C9  & Durability              & Benefit & Portability \& durability \\
    C10 & Battery life            & Benefit & Portability \& durability \\
    C11 & Connectivity ports      & Benefit & Ecosystem utility \\
    C12 & Software compatibility  & Benefit & Ecosystem utility \\
    C13 & Brand                   & Benefit & Trust \& after-sales \\
    C14 & After-sales / warranty  & Benefit & Trust \& after-sales \\
    \bottomrule
  \end{tabularx}
  \floatsource{\emph{Source:} Authors.}
\end{table}

\subsection{Interview-Driven Refinement}\label{subsec:interview-outcomes}

Following the revision rule described in \Cref{subsec:expert-interviews}, the interviews
changed the candidate set in two ways. First, the dedicated graphics card was removed.
Four of the five experts converged independently on the argument that a discrete GPU is a
cost without a corresponding benefit for office work, since it raises the purchase price,
weight, and thermal load while shortening battery life, and the budget it consumes is
better redirected towards a newer-generation processor, more memory, or faster storage.
This judgement is consistent with the finding of \textcite{maghsoudi_hybrid_2026} that the
graphics card ranked last among ten notebook attributes in their expert-weighted model.
Second, software compatibility was added. The experts who work inside Vietnamese offices
(CG-01, CG-03) described a lock-in effect that the international literature rarely makes
explicit, in which accounting packages, internal ERP systems, and the government's customs
declaration software are written for Windows, so a device that cannot run them is excluded
from consideration regardless of its other merits. The interviews also sharpened the
definitions of two retained criteria, restricting the storage criterion to solid-state
drives and extending the design criterion to cover shell material, and recalibrated the
expectation for battery life, because mains power is readily available in offices, so a
runtime of four to seven hours was considered sufficient, a markedly lower bar than the
top-priority status that battery life receives in \textcite{weng_siew_lam_evaluation_2023}.
The revised set therefore still contains fourteen criteria, but with GPU replaced by
software compatibility and with definitions anchored to local working conditions; this is
the candidate set reported in \Cref{tab:criteria} above and carried into the screening
survey of \Cref{subsec:screening}.

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[
      box/.style={draw, rounded corners=2pt, align=center, text width=6.8cm,
                  minimum height=14mm, font=\footnotesize},
      arr/.style={-{Stealth[length=2.5mm]}, thick},
      note/.style={font=\scriptsize, align=left}
    ]
    \node[box] (initial) {Literature-derived set\\14 criteria\\includes GPU};
    \node[box, below=15mm of initial] (revised) {Expert-revised set\\14 criteria\\GPU out; software compatibility in};
    \node[box, below=15mm of revised] (retained) {Survey-retained set\\7 criteria\\mean $\geq 3.0$};
    \draw[arr] (initial) -- node[right, note] {expert refinement\\($n=5$)} (revised);
    \draw[arr] (revised) -- node[right, note] {Likert screening\\($N=84$)} (retained);
  \end{tikzpicture}
  \caption{Transition from the literature-derived criteria to the final retained set.}
  \label{fig:criteria-transition}
  \floatsource{\emph{Source:} Authors.}
\end{figure}

\section{Criteria Screening Results}\label{sec:screening}

The screening survey described in \Cref{subsec:screening} collected 84 responses, all of
which were complete. No respondent left more than 20\% of the items unanswered, and no
straight-lined response was detected, so every questionnaire entered the analysis.
\Cref{tab:survey-results} reports the descriptive statistics for the fourteen candidate
criteria together with the retention decision under the pre-committed threshold of a mean
of at least 3.0.

\input{tables/survey-results}

Seven criteria pass the threshold. Battery life records the highest mean (4.02), followed
by memory and physical size or weight (both 3.96), storage (3.89), processor (3.86), and
price and brand (both 3.79). The seven removed criteria all fall clearly below the
midpoint, from keyboard experience (2.88) down to durability (2.36). The gap between the
lowest retained mean (3.79) and the highest removed mean (2.88) spans almost a full scale
point, so the composition of the retained set is not sensitive to small movements of the
threshold around the midpoint. These seven criteria constitute the answer to RQ1 and
define the dimension of the pairwise comparison in Stage~2. It is worth recording, ahead
of the fuller treatment in \Cref{ch:discussion}, that several criteria emphasised in the
expert interviews, notably software compatibility, durability, and after-sales support,
were screened out by the wider respondent pool, a divergence between the expert panel and
the survey respondents that the two-step design of Stage~1 makes visible rather than
concealing.

\section{Criteria Weights and Consistency Check}\label{sec:weights}

The retained criteria were compared pairwise by the representative expert as described in
\Cref{sec:stage2}, producing a $7 \times 7$ reciprocal matrix from 21 elicited judgements.
Before the weights were interpreted, the matrix was submitted to the consistency check of
\Cref{eq:cr}. The principal eigenvalue is $\lambda_{\max} = 7.3823$, giving a consistency
index of $CI = 0.0637$ and, with the random index $RI = 1.32$ for $n = 7$, a consistency
ratio of $CR = 0.0483$. As this value lies below the acceptance threshold of 0.10
\parencite{Saaty2008DECISIONMW}, the judgement set is internally coherent and the derived
weights are usable without re-elicitation. \Cref{tab:ahp-weights} presents the resulting
priority vector.

\input{tables/ahp-weights}

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[x=14cm,y=0.72cm,font=\footnotesize]
    \foreach \name/\value/\y in {Price/.3491/7,CPU/.2769/6,SSD/.1592/5,RAM/.1201/4,Brand/.0430/3,{Size / weight}/.0258/2,{Battery life}/.0258/1}{
      \fill[gray!65] (0,\y-0.22) rectangle (\value,\y+0.22);
      \node[anchor=east] at (-0.01,\y) {\name};
      \node[anchor=west] at (\value+0.006,\y) {\pgfmathprintnumber[fixed,precision=4]{\value}};
    }
    \draw[->] (0,0.45) -- (0.38,0.45) node[right] {Weight};
    \foreach \x in {0,.1,.2,.3}{\draw (\x,0.40)--(\x,0.50) node[below=2mm] {\pgfmathprintnumber[fixed,precision=1]{\x}};}
  \end{tikzpicture}
  \caption{AHP weights of the seven retained criteria.}
  \label{fig:ahp-weights}
  \floatsource{\emph{Source:} Authors.}
\end{figure}

The weight structure is strongly concentrated at the top. Price carries the largest weight
(0.3491) and the processor the second largest (0.2769); together these two criteria
account for approximately 62.6\% of the total priority, so the model characterises the
office-laptop decision primarily as a trade-off between budget and computing capability.
A middle tier follows, comprising storage (0.1592) and memory (0.1201), while brand
(0.0430), size or weight (0.0258), and battery life (0.0258) form a low-priority tier that
collectively holds less than 10\% of the weight. The position of battery life is notable,
because although it recorded the highest mean in the screening survey, it receives the
smallest AHP weight, and the same reversal separates this result from
\textcite{weng_siew_lam_evaluation_2023}, who report battery life as the most heavily
weighted criterion in their mobile-phone model. This pattern is consistent with the
recalibration that emerged from the expert interviews, in which battery life is a
requirement that most office laptops already satisfy rather than a differentiator; its
interpretation is developed in \Cref{ch:discussion}. As a robustness check, the weights
were recomputed by the geometric-mean method described in \Cref{sec:stage2}; the resulting
vector preserves the same ordering of all seven criteria, indicating that this ordering is
stable across the two computational methods applied here rather than an artefact of the
eigenvector procedure specifically. These weights answer RQ2 and enter Stage~3 as the
weighting vector for TOPSIS.

\section{Decision Scenario and Alternatives}\label{sec:scenario}

Applying the outlet pool and deduplication procedure of \Cref{subsec:alternatives} to the
seven review outlets of \Cref{tab:review-sources} produced twelve distinct candidate
laptops after configuration variants of the same model were collapsed into a single entry.
Price and configuration data for all twelve candidates were collected from Amazon.com on
2026-07-12, following the sourcing rationale of \Cref{subsec:alternatives}. To keep the
ranking table and the calculation tables of \Cref{app:laptop-data} compact, each candidate
is assigned a short identifier, Can1 through Can12, in the order of collection;
\Cref{tab:candidate-ids} reports the assignment, and the subsequent tables refer to the
candidates by these identifiers.

\begin{table}[htbp]
  \centering
  \caption{Candidate identifiers assigned to the twelve laptop models.}
  \label{tab:candidate-ids}
  \begin{tabular}{@{}ll@{}}
    \toprule
    Candidate & Model \\
    \midrule
    Can1  & Lenovo ThinkPad X1 Carbon Gen 13 Aura Edition \\
    Can2  & Lenovo ThinkPad X1 Carbon Gen 14 Aura Edition \\
    Can3  & Apple MacBook Air M4 \\
    Can4  & Apple MacBook Air M5 \\
    Can5  & Asus ExpertBook P5 \\
    Can6  & HP EliteBook Ultra G1i \\
    Can7  & Lenovo ThinkPad X9 15 Aura Edition (2025) \\
    Can8  & Dell 14 Premium (XPS 14 2026) \\
    Can9  & Lenovo ThinkPad T14 Gen 6 (AMD) \\
    Can10 & Lenovo ThinkPad T14 Gen 7 \\
    Can11 & Acer Aspire 5 \\
    Can12 & Lenovo ThinkPad X1 Carbon Gen 12 \\
    \bottomrule
  \end{tabular}
  \floatsource{\emph{Source:} Authors.}
\end{table}

\Cref{tab:decision-matrix} reports the attribute values in the form in which they were
collected, that is, before the CPU and Brand values are converted into the PassMark
benchmark score and the Gartner market-share figure described in
\Cref{subsec:operationalisation}; that converted matrix is the input actually used by the
TOPSIS procedure of \Cref{sec:ranking}.

\input{tables/decision-matrix}

The twelve candidates span a nearly ninefold price range, from \$395 for the
entry-level Acer Aspire 5 to \$3{,}499 for the top-specification ThinkPad X1 Carbon Gen 14
Aura Edition, and are drawn from six brands, with Lenovo contributing six of the twelve
listings. Every candidate except the Acer Aspire 5 (8\,GB) and the Lenovo ThinkPad T14
Gen 6 AMD (16\,GB) carries 32\,GB of RAM or the 24\,GB unified-memory configuration used
by Apple's M-series chips, so the RAM criterion mainly separates these two candidates from
the rest of the field rather than distinguishing among the other ten. The benchmark and
market-share conversion applied to the CPU and Brand columns, together with the full
step-by-step TOPSIS calculation, is reported in \Cref{app:laptop-data}.

\section{TOPSIS Ranking Results}\label{sec:ranking}

Applying the seven-step TOPSIS procedure of \Cref{subsec:topsis-procedure} to the converted
decision matrix of \Cref{app:laptop-data-conversion}, using the AHP weights of
\Cref{tab:ahp-weights}, produces the closeness coefficients and ranking reported in
\Cref{tab:topsis-ranking}. The full normalisation, weighting, and ideal-solution
calculations underlying this table are reported step by step in \Cref{app:laptop-data-topsis}.

\input{tables/topsis-ranking}

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[x=12cm,y=0.52cm,font=\scriptsize]
    \foreach \id/\value/\y in {Can9/.6152/12,Can8/.6046/11,Can4/.6039/10,Can6/.5901/9,Can3/.5868/8,Can5/.5621/7,Can11/.5608/6,Can12/.5444/5,Can7/.5416/4,Can1/.5170/3,Can10/.4140/2,Can2/.3803/1}{
      \draw[gray!45] (0.35,\y) -- (\value,\y);
      \fill (\value,\y) circle (1.6pt);
      \node[anchor=east] at (0.34,\y) {\id};
      \node[anchor=west] at (\value+0.006,\y) {\pgfmathprintnumber[fixed,precision=4]{\value}};
    }
    \draw[->] (0.35,0.25) -- (0.66,0.25) node[right] {$C_i$};
    \foreach \x in {.4,.5,.6}{\draw (\x,0.20)--(\x,0.30) node[below=2mm] {\pgfmathprintnumber[fixed,precision=1]{\x}};}
  \end{tikzpicture}
  \caption{TOPSIS closeness coefficients of the twelve candidate laptops.}
  \label{fig:topsis-coefficients}
  \floatsource{\emph{Source:} Authors.}
\end{figure}

The Lenovo ThinkPad T14 Gen 6 (AMD) obtains the highest closeness coefficient
($C_i = 0.6152$), ahead of the Dell 14 Premium ($C_i = 0.6046$) and the Apple MacBook Air M5
($C_i = 0.6039$). The three leading alternatives are separated by no more than 0.011 in
closeness coefficient, a narrow margin that suggests the top of the ranking may be
sensitive to small changes in the underlying scores, though this was not tested through a
formal sensitivity analysis. The top-ranked model reaches this position without dominating
any single criterion outright. Its price of \$1{,}100 is the second lowest in the set, its
processor score (19{,}546) is in the middle of the field, and its 16\,GB of RAM is the
smallest configuration among the eleven non-Acer candidates; it does share the
joint-highest Brand score (26.5) with several other Lenovo entries in the set, but this
shared strength combines with a below-median price rather than a lead on any criterion by
itself, and no single weakness is severe enough to offset that combination. The Dell 14
Premium and the MacBook Air M5 reach second and third place from a different combination,
in which both carry the highest or near-highest processor scores in the field (30{,}684
and 26{,}831 respectively), which compensates for a price above the sample median.

The lowest-ranked alternative is the Lenovo ThinkPad X1 Carbon Gen 14 Aura Edition
($C_i = 0.3803$), the most expensive candidate at \$3{,}499, despite a mid-to-upper
processor score (20{,}946, the fourth highest in the set); because Price and CPU together
account for 62.6\% of the total criteria weight (\Cref{sec:weights}), even a comfortably
above-average score on one of these two criteria does not compensate for an extreme value
on the other. The Acer Aspire 5, the cheapest candidate at \$395, ranks seventh of twelve
($C_i = 0.5608$) rather than first, even though Price carries the single largest weight in
the model, because its processor, memory, storage, and brand-share values are the lowest
or near-lowest on every one of those criteria, and the combined deficit outweighs its
price advantage. Taken together, the two
extremes illustrate that no criterion, however heavily weighted, guarantees a favourable
rank on its own; TOPSIS rewards a candidate's overall balance across the weighted profile
rather than dominance on a single attribute. The practical implication of this result,
together with its bearing on the recommended alternative, is developed further in
\Cref{ch:discussion}.
