\chapter{Discussion}\label{ch:discussion}

% ============================================================
% Cau truc hoc tu paper Q1: Maghsoudi 2026 (Discussion + Managerial Insights rieng),
% Saputro 2023 (muc 5.4 Managerial implication rieng).
% Diem ban cua de tai nam o 5.1: doi chieu ly thuyet quoc te vs thuc tien VN.
% ============================================================

\section{Interpretation of the Criteria Weights}\label{sec:interp-weights}

The weight structure reported in \Cref{sec:weights} concentrates 62.6\% of the total
priority in price (0.3491) and processor capability (0.2769), and this concentration is
readable directly from the decision context the criteria were calibrated against. In the
expert interviews, the purchase budget for an office machine is not a preference to be
optimised but a policy set in advance by management or the IT department, so price operates
as the framing variable within which every subsequent hardware trade-off is made. The
dominant position of price is consistent with the wider consumer-behaviour evidence:
\textcite{liao_determinants_2022} find price the single most influential attribute in their
conjoint analysis of laptop choice, and \textcite{sostar_assessment_2023} identify personal
budget as the strongest driver among the cultural, social, personal, and psychological
factors they weigh. The weight of the processor, in turn, reflects the interviews'
consistent redirection of scarce budget towards computing capability: the experts' argument
for removing the dedicated graphics card (\Cref{subsec:interview-outcomes}) was precisely
that the money it consumes buys more useful performance when spent on a newer-generation
CPU, more memory, or faster storage.

The more interesting feature of the weight vector, however, is its bottom tier, because
this is where the local calibration departs most visibly from the international
literature. Battery life records the highest mean in the screening survey (4.02,
\Cref{sec:screening}) yet receives the joint-smallest AHP weight (0.0258,
\Cref{sec:weights}), and \textcite{weng_siew_lam_evaluation_2023} report battery life as
the most heavily weighted criterion in their integrated AHP--TOPSIS model. The apparent
contradiction dissolves once the two instruments are read as measuring different things.
The Likert screening asks whether a criterion matters at all, and battery life clearly
does: a machine that cannot hold a charge through a meeting is unusable. The pairwise
comparison instead asks how much of a decisive difference the criterion makes between
candidate machines, and here the expert judgement recorded in the interviews applies:
mains power is readily available in Vietnamese offices, a runtime of four to seven hours
already covers meetings and short mobile work, and most current office laptops meet that
bar. Battery life therefore behaves as a threshold requirement, broadly satisfied across
the candidate set, rather than as a differentiator, and a rational pairwise comparison
assigns differentiating weight elsewhere. The same logic extends to physical size and
weight, which the survey ranks joint-second in importance (3.96) but the AHP places in the
same low tier (0.0258): the interviews emphasised weight because motorbike commuting makes
a heavy machine a daily burden, yet the office segment has converged on 14-inch machines
weighing roughly 1.3 to 1.6 kilograms, so the attribute varies too little across realistic
candidates to move a pairwise judgement. What the survey measures as salience, the AHP
re-expresses as marginal decision value, and criteria whose requirements the market
already meets lose weight in the second instrument without losing importance in the first.

A related divergence between the expert and end-user perspectives concerns the criteria
that the screening survey removed. Software compatibility was added to the candidate set
on the strength of the interviews, in which office workers described a Windows lock-in
effect created by accounting packages, internal ERP systems, and the government's customs
declaration software (\Cref{subsec:interview-outcomes}), yet the wider respondent pool
screened it out, together with durability and after-sales support
(\Cref{sec:screening}). The likely mechanism is that end users experience compatibility as
a background condition rather than a choice variable: almost every machine they encounter
in a Vietnamese office already runs Windows, so the criterion only becomes visible at the
moment a non-Windows device enters consideration, which is exactly the procurement-risk
scenario the experts are paid to anticipate. The two-step design of Stage~1 makes this
divergence observable instead of averaging it away, and the practical reading is that the
screened-out criteria function as eligibility conditions applied before the weighted
comparison, not as attributes the weighted comparison should ignore.

One divergence worth flagging concerns the granularity of the Display criterion, which the
screening survey of \Cref{sec:screening} removed for falling just below the retention
threshold (mean $=2.80$, $SD=1.32$). Read against the wider literature, this outcome may
reflect how the criterion was operationalised rather than a genuine indifference to visual
quality among Vietnamese office workers. \textcite{sonmez_cakir_determination_2020}
decompose monitor properties into four distinct sub-criteria, resolution, size,
touch-screen support, and other usage advantages, and find that these carry markedly
different weights within the group, with touch-screen support outweighing resolution.
\textcite{weng_siew_lam_evaluation_2023} go further and place screen size and pixel density
in two separate higher-level categories, physical properties and technical specification
respectively, with screen size ranking second within its category while pixel density ranks
last within its own; the two attributes therefore answer different buyer concerns, physical
footprint versus visual output quality, rather than a single underlying preference.
\textcite{maghsoudi_hybrid_2026} report the starkest contrast: in their sentiment-weighted
ranking of sixteen notebook features, display quality places third overall, immediately
behind price and design, while size places eleventh, a gap that would be invisible had the
two been folded into one composite score. Set beside these findings, the single ``Display''
item used in the present screening instrument bundled resolution, sharpness, anti-glare
coating, and brightness into one rating, plausibly diluting a genuinely important
sub-attribute, visual quality, with a less decisive one, raw screen size, and depressing the
mean below the retention threshold. This is offered as an interpretive caveat rather than a
claim that Vietnamese office workers disregard display quality outright, and it points to
criterion granularity, not only criterion choice, as a methodological lever meriting
attention in future replications of this instrument.

\section{Interpretation of the Ranking}\label{sec:interp-ranking}

The identity of the top-ranked alternative is itself informative about what the calibrated
model rewards. The Lenovo ThinkPad T14 Gen 6 (AMD) leads the ranking
(\Cref{sec:ranking}) without holding the best value on any single criterion: its price is
the second lowest in the set, its processor sits in the middle of the field, and its
16\,GB of memory is the smallest configuration among the eleven non-Acer candidates. Its
closeness coefficient of 0.6152 is instead the product of a profile with no severe
weakness, anchored by a below-median price and supported by the largest brand market
share in the set. This is precisely the machine the expert interviews described before
any market data had been collected: a current-generation mid-tier processor was judged
sufficient for administrative workloads, and the budget saved by not chasing peak
specifications was to be preserved rather than redirected into features the office
segment does not use. That the data-driven ranking independently converges on this
profile is a meaningful internal consistency check between Stage~1's qualitative
calibration and Stage~3's quantitative outcome.

The two extremes of the ranking reinforce the same reading from opposite directions. The
Acer Aspire 5, the cheapest candidate at \$395, finishes seventh rather than first despite
price carrying the single largest weight, because its deficits on processor, memory,
storage, and brand share accumulate faster than its price advantage compensates. At the
other end, the ThinkPad X1 Carbon Gen 14 Aura Edition finishes last at \$3{,}499 despite
the second-highest processor score in the set, an outcome that echoes the
over-specification argument of \textcite{rau_optimal_2018}: paying a premium for capability
beyond the requirements of the use case destroys rather than creates value once price
enters the comparison with its full weight. Between these extremes, the top three
alternatives are separated by less than 0.011 in closeness coefficient, so the ranking is
better read as producing a short list of near-equivalent recommendations, the T14 Gen 6,
the Dell 14 Premium, and the MacBook Air M5, than as declaring a single decisive winner;
the Dell and the Apple machine reach their positions through the complementary route of
field-leading processor scores that offset above-median prices. Within that short list,
the practical tie-breakers are exactly the eligibility conditions identified in
\Cref{sec:interp-weights}: the MacBook's macOS platform, for instance, would be filtered
out by the Windows lock-in constraint in offices where that constraint binds, even though
the weighted comparison itself ranks it third.

\section{Practical and Managerial Implications}\label{sec:implications}

For an individual office worker buying with personal funds, the model's main practical
export is not the specific ranking, which is tied to a catalogue snapshot, but the
structure of the weights: roughly 63\% of the decision is a price--processor trade-off,
storage and memory form a middle tier worth attending to, and battery life, weight, and
brand deserve verification against a threshold rather than optimisation. Concretely, the
buyer should fix the budget first, secure a current-generation mid-tier processor and
16\,GB of memory within it, confirm that battery life and weight fall in the acceptable
band for their commuting pattern, and resist paying premiums for capability at the top of
the range, which the ranking shows to be counterproductive under office-work weights.

For corporate procurement, the value of the model lies in making an implicit decision
auditable. The interviews indicated that purchase budgets are set by policy and that IT
managers already think in terms of total cost of ownership; the AHP--TOPSIS pipeline
turns those practices into an explicit weight vector and a reproducible ranking that can
be attached to a purchase justification, revisited when the catalogue changes, and
recalculated by replacing a single input matrix. Because the weights and the alternatives
are decoupled, a procurement office can hold the weight vector stable across purchasing
rounds while refreshing the decision matrix, which converts a one-off analysis into a
reusable procedure. The screened-out criteria retain a role here as eligibility gates:
software compatibility and after-sales coverage are best enforced as pass-or-fail
conditions before the weighted comparison, mirroring how the experts described actual
procurement practice.

The results also carry signals for the supply side. For retailers advising office
customers, the ranking suggests that the persuasive assortment is not the cheapest machine
or the most powerful one but the balanced mid-tier configuration, and that advice framed
as ``sufficient processor within budget'' matches how this segment actually weighs the
purchase. For manufacturers, the weight structure implies that marginal engineering
investment aimed at this segment earns more return in price positioning and
processor-generation currency than in pushing battery life or thinness beyond the
threshold the segment already considers satisfied; a battery lasting beyond the four-to-seven-hour
working band, in particular, adds cost against a criterion holding under 3\% of
the decision weight.

\section{Comparison with Prior Studies}\label{sec:comparison}

Set against the prior studies reviewed in \Cref{sec:prior-studies}, the present results
agree on the centre of the priority structure and diverge at its edges, and the
divergences are traceable to context rather than method. The dominance of price is the
most stable point of agreement: it matches the conjoint evidence of
\textcite{liao_determinants_2022}, the budget-driven consumer behaviour reported by
\textcite{sostar_assessment_2023}, and the price-sensitive student segment of
\textcite{elnatan_alternatif_nodate}, in which price, processor, and memory jointly lead
the weights, a composition strikingly close to the top of the present vector.

The clearest divergence concerns brand. \textcite{sonmez_cakir_determination_2020} find
brand image the highest-weighted criterion among their six, whereas the present model
assigns brand only 0.0430. The difference is explicable through the signalling mechanism
that motivates brand weight in the first place: brand compensates for information
asymmetry when buyers cannot evaluate technical quality directly, and that asymmetry is
largest for general consumers choosing alone. In the office-procurement context studied
here, IT managers and standardised specifications absorb much of that uncertainty, so the
signalling value of brand shrinks even while brand retains a residual role, visible in the
ranking, where Lenovo's market share contributes to the winning profile. The second
divergence, the position of battery life relative to
\textcite{weng_siew_lam_evaluation_2023}, was examined in \Cref{sec:interp-weights} and
reflects both the device category, since a phone cannot assume mains power, and the
office setting that can.

At the level of ranked outcomes, comparison is necessarily looser because candidate sets
differ, but two observations hold. \textcite{gaur_application_2023}, ranking laptop brands
during the pandemic with a comparable AHP--TOPSIS apparatus, identify Dell as the leading
alternative; the present ranking, built on model-level rather than brand-level
alternatives, places a Dell machine second, suggesting that the earlier result survives
translation to a finer-grained alternative space. \textcite{maghsoudi_hybrid_2026},
weighting notebook features by mined sentiment, place the graphics card last among
sixteen features and price near the top, which independently corroborates both the
experts' removal of the GPU from the candidate set and the weight the present model gives
to price; their finding that display quality ranks third, against its removal by the
present screening survey, is the one substantive disagreement, and
\Cref{sec:interp-weights} argues it is most plausibly an artefact of criterion
granularity rather than a genuine difference in what office users value. Taken together,
the comparisons support the position claimed in \Cref{sec:gap}: the international
literature travels well at the core of the priority structure, price and processing
capability, and requires local recalibration precisely at the criteria, battery, weight,
brand, and software ecosystem, whose value depends on infrastructure and working
conditions.
